\documentclass[16pt]{article}
\usepackage[english]{babel}
\usepackage{longtable}
\usepackage[top=1in, bottom=0.25in, left=1.25in, right=1.25in,includefoot,heightrounded]{geometry}
\usepackage{indentfirst}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,tikz}
\usepackage{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[document]{ragged2e}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{listings}
\usepackage{color}
\usepackage{flowchart}
\usepackage{hyperref}
\usepackage{graphicx}
\usetikzlibrary{arrows}


\usetikzlibrary{shapes.geometric, arrows}
\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
\tikzstyle{decision} = [diamond, minimum width=4cm, minimum height=0.5cm, text centered, draw=black, fill=green!30]
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
\tikzstyle{arrow} = [thick,->,>=stealth]
\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=2cm, text width=4cm, minimum height=1cm, text centered, draw=black, fill=blue!30]

\pagestyle{fancy}
\fancyhf{}
\lhead{Myles Deslippe}
\rhead{Comp 3670 | Computer Networks}
\cfoot{\thepage}

\definecolor{MyDarkGreen}{rgb}{0.0,0.4,0.0}
\lstset{inputencoding=ansinew}
\lstset{breaklines=true} 

\begin{document}

    \section*{\centering{Transport Services and Protocols}}

    \subsection*{Transport Services and Protocols}
    \begin{itemize}
        \item The \textbf{transport layer} is responsible for providing \textbf{logical communication} between \textbf{application procceses} running on \textbf{different hosts}.
        \item There are two actions the transport layer needs to achieve:
        \begin{enumerate}
            \item \textbf{Send action}: The transport layer needs to \textbf{divide application messages into segments}, and pass them to the \textbf{network layer}.
            \item \textbf{Recieve action}: The transport layer needs to \textbf{reassemble segments} from the \textbf{network layer} into \textbf{messages}, and pass them to the \textbf{application layer}.
        \end{enumerate}
        \item There are \textbf{two transport protocols} available: the \textbf{User Datagram Protocol (UDP)}, and the \textbf{Transmission Control Protocol (TCP)}.
    \end{itemize}

    \subsection*{The Application Layer, The Transport Layer, and The Network Layer}
    \begin{itemize}
        \item The \textbf{application layer} is the \textbf{process} that is running on the \textbf{network host}.
        \item The \textbf{transport layer} is responsible for the \textbf{logical communcation} between \textbf{processes}.
        \item The \textbf{network layer} is responseible for the \textbf{logical communication} between \textbf{network hosts}.
        \item When a \textbf{message is sent} from an application over the network, it follow the following order: application layer $\rightarrow$ transport layer $\rightarrow$ network layer.
        \item When a \textbf{message is received} from the network, and sent to an application, it follows the following order: network layer $\rightarrow$ transport layer $\rightarrow$ application layer.
    \end{itemize}

    \subsection*{Transport Layer Actions}
    \begin{itemize}
        \item When a message is \textbf{sent from an application}, the \textbf{transport layer} does the following:
        \begin{enumerate}
            \item Recieves the application-layer message.
            \item Determines segment header field values.
            \item Creates the segment.
            \item Passes the segment to the internet protocol.
        \end{enumerate}
        \item When a message is \textbf{recieved from the network}, the \textbf{transport layer} does the following:
        \begin{enumerate}
            \item Recieves the segment from the internet protocol.
            \item Checks the header values.
            \item Extracts the application-layer message.
            \item Demultiplexes message up to the application via a socket.
        \end{enumerate}
    \end{itemize}

    \subsection*{User Datagram Protocol and Transmission Control Protocol}
    \begin{itemize}
        \item The \textbf{User Datagram Protocol (UDP)} is an \textbf{unreliable}, \textbf{unordered} delivery protocol. This protocol has \textbf{minimal overhead} but is \textbf{not reliable}.
        \item The \textbf{Transmission Control Protocol (TCP)} is a \textbf{reliable}, \textbf{in-order delivery} protocol that supports \textbf{congestion control}, \textbf{flow control}, and \textbf{connection setup}. This protocol has \textbf{significant overhead}, but is \textbf{very reliable}.
        \item Both \textbf{UDP} and \textbf{TCP} do not provide \textbf{delay guarantees} and \textbf{bandwith guarantees}.
    \end{itemize}

    \section*{\centering{Multiplexing}}
    
    \subsection*{Multiplexing}
    \begin{itemize}
        \item \textbf{Multiplexing} is a method used by \textbf{networks} to \textbf{consolidate multiple signals} into a \textbf{single composit signal} that is then \textbf{transported over a common medium}. 
        \item When \textbf{sending data}, \textbf{multiplexing} is used to \textbf{transmit segments} from \textbf{several sockets} via \textbf{transport headers}.
        \item When \textbf{recieving data}, the \textbf{header info} is used to \textbf{demultiplex} the data, sending it to the \textbf{correct socket}.
        \item \textbf{Demultiplexing} works by receiving \textbf{IP datagrams} containing a \textbf{source and destination IP address and port as headers}. Each datagram contains a \textbf{single segment}.
        \begin{itemize}
            \item The IP addresses and port numbers are used to route the segment to the correct socket.
            \item This is why you must specify a port number when creating a socket.
        \end{itemize} 
    \end{itemize}

\end{document}