\documentclass{article}
\usepackage[english]{babel}
\usepackage{longtable}
\usepackage[top=1in, bottom=0.25in, left=1.25in, right=1.25in,includefoot,heightrounded]{geometry}
\usepackage{indentfirst}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,tikz}
\usepackage{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[document]{ragged2e}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{listings}
\usepackage{color}
\usepackage{flowchart}
\usepackage{hyperref}
\usepackage{graphicx}
\usetikzlibrary{arrows}

\usetikzlibrary{shapes.geometric, arrows}
\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
\tikzstyle{decision} = [diamond, minimum width=4cm, minimum height=0.5cm, text centered, draw=black, fill=green!30]
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
\tikzstyle{arrow} = [thick,->,>=stealth]
\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=2cm, text width=4cm, minimum height=1cm, text centered, draw=black, fill=blue!30]

\pagestyle{fancy}
\fancyhf{}
\lhead{Myles Deslippe}
\rhead{Phil 2210 | Introduction to Ethics}
\cfoot{\thepage}

\definecolor{MyDarkGreen}{rgb}{0.0,0.4,0.0}
\lstset{inputencoding=ansinew}
\lstset{breaklines=true} 

\begin{document}

    \section*{\centering{Morality and Religion}}

    \subsection*{Attempts to Ground Morality in Religion}
    \begin{itemize}
        \item There is a widely held view that there is an \textbf{intimate connection} between \textbf{morality and religion}.
        \begin{itemize}
            \item This implies that if there is no god, then there is no morality.
        \end{itemize}
        \item Even if there is a connection between \textbf{morality and religion}, attempts to \textbf{understand} this connection so far have \textbf{failed}.
        \item Two \textbf{popular} attempts to \textbf{understand the connection} between \textbf{morality and religion} are \textbf{Divine Command Theory}, and \textbf{Natural Law Theory.}
    \end{itemize}

    \subsection*{Divine Command Theory (DCT)}
    \begin{itemize}
        \item The \textbf{roots} of \textbf{DCT} have their origin in the writings of \textbf{Plato} in a dialogue called \textbf{The Euthyphro}.
        \item The basic ideas of \textbf{DCT} can be summarized as follows: 
        \begin{enumerate}
            \item An act is \textbf{obligatory} if God commands it (tells you to do it).
            \item An act is \textbf{permissible} if God permits it.
            \item An act is \textbf{forbidden} if God forbids it (tells you to not do it).
        \end{enumerate}
        \item This raises the question: "How do you know what God commands?"
        \begin{itemize}
            \item Some people believe that it is found in \textbf{religious scripture}.
        \end{itemize}
        \item An \textbf{ontological problem} (a problem in the study of existance) that relates to \textbf{DCT} is: how do we know that God exists?
        \begin{itemize}
            \item There is an ontological proof that goes as follows: God's essence is perfection. It is more perfect to exist than to not exist. Thus, God exists.
            \item However, not everyone accepts theological proofs of God's existance.
        \end{itemize}
        \item Some defenders of \textbf{DCT} believe that it is \textbf{not necessary} to \textbf{believe in God} to adhere to \textbf{DCT}. They would instead argue that \textbf{only if God exists}, morality is possible. So they are giving an account of what morality would be if there were a God.
        \item An \textbf{epistemological problem} (a problem about how we know something) with \textbf{DCT} is: even if God exists, how do we \textbf{know} what \textbf{God's commands are}?
        \begin{itemize}
            \item Some would argue that we know through \textbf{scripture, prayer, meditation, and religious experiences}.
            \item Adams claims that all \textbf{wrong actions} have one thing in common; they are \textbf{contrary} to the \textbf{commands} of a \textbf{loving God}.
        \end{itemize}
    \end{itemize}

    \subsection*{The Major Problem with DCT}
    \begin{itemize}
        \item The \textbf{major problem} with \textbf{DCT} was put forward by \textbf{Plato} in \textbf{Euthyphro}. The \textbf{problem} is known as the \textbf{Euthyprho dilemma}.
        \item The \textbf{Euthyprho dilemma is as follows}: Does god command an action because it is right, or is the action right because God commands it?
        \begin{itemize}
            \item Suppose an action X is right, simply because God commands it. Doesn't that make morality arbitrary? It can be whatver God wants? He can change the commands at any time. God could command us to do something that we think is immoral.
            \item Suppose an action X is right, then God will command it. But this means, that religion is based on morality instead of morality being based on religion. God is essentially acting as a messanger for something above him.
        \end{itemize}
        \item This dilemma leads us to \textbf{abandon DCT}. Though there have been attempts to resuscitate it.
        \begin{itemize}
            \item One theory that attempted to save DCT is \textbf{A New Device Command Theory} by \textbf{Robert Adams}. To get around the \textbf{Euthrphro dilemma} it modifies \textbf{DCT} as follows: An act is morally right if is commanded by a \textbf{loving God}; a loving god would never command us to do something that is morally wrong by our moral intuitions.
        \end{itemize}
    \end{itemize}    

    \subsection*{Natural Law Theory (NLT)}
    \begin{itemize}
        \item Since not every accepts solutions proposed to the \textbf{Euthyphro dilemma}, \textbf{St. Thomas Aquinas} proposed an alternative way of grounding morality in religion. This was is \textbf{Natural Law Theory}.
        \item \textbf{Natural Law Theory} can be summarized in terms of three theses:
        \begin{itemize}
            \item Every event and every action has a \textbf{purpose}.
            \item There are \textbf{laws of nature} that represent the \textbf{purposes} of actions and events.
            \item God provides us with the \textbf{ability to reason} in order to comprehend the \textbf{laws of nature}.
        \end{itemize}
        \item In \textbf{NLT}, moral laws are \textbf{laws of nature}, not \textbf{commands of God}. God enters into morality only at the level of \textbf{providing us with the ability to reason}.
        \begin{itemize}
            \item This allows NLT to avoid the problems that exist with DCT.
        \end{itemize}
        \item Essentially, \textbf{NLT} is a theory that it is \textbf{immoral to use things in a way that does not fit their purpose}.
        \begin{itemize}
            \item An example would be: arms were created to life things, using them for any other purpose (such as to hit someone) is morally wrong.
        \end{itemize}
    \end{itemize}

    \subsection*{Problems with NLT}
    \begin{itemize}
        \item One problem with \textbf{NLT} is that it is not clear that \textbf{every event has a purpose}.
        \begin{itemize}
            \item For example, we do not have eyes for the purpose of seeing, rather we have eyes as an evolutionary adaptation.
        \end{itemize}
        \item Another problem with \textbf{NLT} is that it commits the \textbf{naturalistic fallacy} (infering "should" from "is"). We are deducing moral oughts from "the purpose an event is". 
        \item Arguably the most serious object is that an \textbf{atheist} could still \textbf{act morally} since having the ability to reason \textbf{does not depend} on \textbf{believing in God}.
        \begin{itemize}
            \item Implying God is irrelevant, since one could be moral without believing in God.
        \end{itemize} 
    \end{itemize}

    \subsection*{Moral Codes}
    \begin{itemize}
        \item Another approach to ground morality in religion, discuessed by \textbf{James Rachels}, is that \textbf{religious scriptures can be regarded as moral codes}.
        \item To determine if an act is morally right or morally wrong, you consult the scripture.
        \item The problem with this approach is that \textbf{scripture is open to interpretation}. You could interperate it \textbf{literally, metaphorically, historically, etc}. People can also take scripture passages out of context, or ignore the meaning of the terms at that time in history.
    \end{itemize}

    \section*{\centering{Non-Religious Natural Law Theories}}

    \subsection*{John Finnis' Natural Law}
    \begin{itemize}
        \item According to \textbf{John Finnis}, natural law theory is used by philosophers to make sense of the following claims that seem to be beyond or prior to positivistic law:
        \begin{enumerate}
            \item That was a crime against humanity.
            \item That action was morally unreasonable, inhuman, curel, etc.
            \item That person should be immune for legal liability.
        \end{enumerate}
        \item As \textbf{Finnis} notes, it's almost as if there is an \textbf{implicit appeal} to a \textbf{"higher law"} - a law of nature.
        \item \textbf{Natural law} tries to make sense of these \textbf{claims} by providing a \textbf{general account} of \textbf{good and evil} in human life.
        \item Finnis claims that a general theory of good and evil will contain the following kinds of propositions:
        \begin{enumerate}
            \item \textbf{Normative claims} about what is right and wrong,
            \item \textbf{Non-Normative claims} justifications of the normative claims. 
        \end{enumerate}
        \item Finnis wants to make it clear that \textbf{natural laws} are \textbf{not mysterious entities that can be read off from nature}, and that \textbf{natural law theories} do not claim that ther laws of nature are \textbf{physical laws} (morality $\ne$ science).
        \begin{itemize}
            \item It is important to recognize that natural law theories are distinct from logic, technology, and science.
        \end{itemize}
        \item 
    \end{itemize}

    \subsection*{John Finnis' Natural Law Definition}
    \begin{itemize}
        \item Finnis claims that the term \textbf{law} does \textbf{not mean an imperative / command} by a superior will such as \textbf{God}.
        \item The term \textbf{law} refers to \textbf{standards of action} which are \textbf{normative}, such that \textbf{not following} these \textbf{standards} is \textbf{unreasonable}.
        \item Finnis defines the term \textbf{natural law} as the following:
        \begin{enumerate}
            \item \textbf{Natural law} is not positive | that is, it is prior to any positioning of any individual, group, or legal system.
            \item \textbf{Natural laws} are standards that \textbf{transcend positive law} | that can be used to critically assess positive law.
            \item \textbf{Natural laws} are standards that are objective in the sense that if one does not follow them, the one is in error.
            \item \textbf{Natural laws} promote human fluorishing.
        \end{enumerate}
        \begin{itemize}
            \item Essentially he is saying natural laws don't come from God, they don't necessarily come from us, they are reasonable, they are objective, they shouldn't be confused with laws created by humans, and they promote human fluorishing.
            \item Natural laws are independent of human laws.
        \end{itemize}
        \item 
    \end{itemize}

    \subsection*{John Finnis' Critique of Morala Skepticism}
    \begin{itemize}
        \item Finnis wants to claim that \textbf{all skeptical attempts to criticize natural law theory} are based on what Hume and G.E. Moore called the \textbf{naturalistic fallacy} (the fallacy of trying to derive "should" from "is").
        \item He wants to \textbf{distinguish non-skeptical theories} from \textbf{natural law theory}, i.e. not all ethical theories that avoid the naturalistic fallacy are necessarily natural law theories. He does this with the following claims:
        \begin{enumerate}
            \item Natural law theories vs other theories are willing to identify certain universal human goods that contribute to human flourishing, such as knowledge, life, friendship.
            \item Natural law theories do not rest on the assumption that moral truths are apprehended by intuition, or that they can be read off from nature. Rather, moral truths are based on "data" that involve theoretical knowledge and true opinion.
            \item Natural law theories do not posit an overarching end to which all human actors refer (for example happiness with Mill).
            \item Natural law theories also try to clarify the methods of the social sciences.
        \end{enumerate}
    \end{itemize}
    
    \subsection*{Ross Shafer-Landau's Natural Law}
    \begin{itemize}
        \item Ross begins by noting that we \textbf{we are all animals}; we are natural creatures.
        \item Like \textbf{all natural creatures}, we need \textbf{food, shelter,} and are \textbf{vulnerable to harm}.
        \item Ross believes that since we are \textbf{a part of nature}, \textbf{morality} may lie in \textbf{knowing our place} in the \textbf{natural order} of things.
        \item Ross  claims that this can be done by doing the best thing for human nature. But that just creates another question: what is human nature?
        \item He discusses two competing conceptions of human nature, but ultimately rejects both:
        \begin{enumerate}
            \item Human nature is what is \textbf{innately human} (not thins we acquire through socialization, learning, etc).
            \begin{itemize}
                \item He goes on to reject this conception on the grounds that some innate traits humans have are repulsive (such as greed, or mistrust).
                \item He reaches the conclusion that \textbf{the morality of our actions does not depend on whether they are innate or acquired}.
            \end{itemize}
            \item Human nature is what \textbf{all huams have in common} (implying unique traits are accidental or non-natural).
            \begin{itemize}
                \item He goes on to reject this conception on the grounds that there may be no universal human characteristics, and even if there are, they may not be able to provide moral guidance.
            \end{itemize}
        \end{enumerate}
        \item He goes on to give the following definition: \textbf{Human nature} is what we are designed \textbf{to be}, and \textbf{to do}. It is some functions of ours, some purpose that we are ment to serve, and some end that we were designed for.
        \item He grants that \textbf{each organ} we have, have a \textbf{purpose}, but wants to find out if it \textbf{makes sense} to say that \textbf{the person} as a whole \textbf{has a function}.
        \begin{itemize}
            \item Some people call this the \textbf{fallacy of composition}; when you assume that a property that holds true for a part of something, is true for the entire thing.
            \item An example would be assuming that NaCl is dangerous because Na is highly reactive, and Cl is toxic.
        \end{itemize}
        \item He proposes two attempts to make sense of his concept of \textbf{natural purpose}:
        \begin{enumerate}
            \item The \textbf{Efficiency Model of Natural Purpose}: According to the efficiency model, we can determine a thing's \textbf{natural purpose} by \textbf{seeing what it is most efficient at doing}. Thus, the natural purpose of humans is determined by what they are most efficient at doing. Humans are efficient at writing essays, solving puzzles, building things, etc.
            \begin{itemize}
                \item There is a problem with this model, if some humans are not good at writing essays or solving puzzles, that implies they are immoral; which is absurd.
                \item Also we are efficient at building weapons and killing people, which is clearly not moral.
            \end{itemize} 
            \item The \textbf{Fitness Model of Natural Purpose}: One could argue that the \textbf{natural purpose} of \textbf{humans} is to \textbf{survive and reproduce}, allowing them to pass on their genes to future generations. Actions that make improve one's fitness (by helping them pass on their genes) are morally right, anything else is morally wrong.
            \begin{itemize}
                \item The problem with this mode, is that immoral things can improve our fitness, such a stealing, killing, etc. Thus this is not a good model.
            \end{itemize}
        \end{enumerate}
    \end{itemize}

\end{document}