\documentclass[16pt]{article}
\usepackage[english]{babel}
\usepackage{longtable}
\usepackage[top=1in, bottom=0.25in, left=1.25in, right=1.25in,includefoot,heightrounded]{geometry}
\usepackage{indentfirst}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,tikz}
\usepackage{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[document]{ragged2e}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{listings}
\usepackage{color}
\usepackage{flowchart}
\usepackage{hyperref}
\usepackage{graphicx}
\usetikzlibrary{arrows}

\usetikzlibrary{shapes.geometric, arrows}
\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
\tikzstyle{decision} = [diamond, minimum width=4cm, minimum height=0.5cm, text centered, draw=black, fill=green!30]
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
\tikzstyle{arrow} = [thick,->,>=stealth]
\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=2cm, text width=4cm, minimum height=1cm, text centered, draw=black, fill=blue!30]

\pagestyle{fancy}
\fancyhf{}
\lhead{Myles Deslippe}
\rhead{Phil 2210 | Introduction to Ethics}
\cfoot{\thepage}

\definecolor{MyDarkGreen}{rgb}{0.0,0.4,0.0}
\lstset{inputencoding=ansinew}
\lstset{breaklines=true} 

\begin{document}

    \section*{\centering{Moral Development}}

    \subsection*{Kohlberg's Theory of Moral Development}
    \begin{itemize}
        \item William Kohlberg wanted to extend Piaget's theory of development to moral development.
        \item According to Kohlberg, there are \textbf{six stages of moral development}:
        \begin{enumerate}
            \item The child is \textbf{entirely selfish}, and does not recognize the needs of others (psychological egosim).
            \item The child \textbf{learns} that by \textbf{begin kind} to others, there will be reciprociy (ethical egoism).
            \item The individual sees the \textbf{family} as the significant moral unit, such that what is morally good, is what is good for the family (virtue ethics).
            \item The individual sees the \textbf{nation and laws} as the significant moral unit, so that what is morally good, is good for the nation (social contract theory).
            \item The individual begins to \textbf{question all of their values}, abandoning both the family and the law (cultural relativism).
            \item The individual begins to \textbf{regard moral laws as being universally valid} (deontology). 
        \end{enumerate}
        \begin{itemize}
            \item The first two stages are said to be \textbf{pre-conventional stages of moral development}.
            \item The middle two stages are said to be \textbf{conventional stages of moral devlopment}.
            \item The last two stages are said to be \textbf{post-conventional stages of moral development}.
            \item Kohl regards \textbf{stage six} as being the \textbf{highest stage of moral development}.
            \item Kohl also claimed that \textbf{women} only make it to \textbf{stage 3} since they \textbf{tend to stay at home} (this was before the women's liberation, when men ran everything).
        \end{itemize}
    \end{itemize}

    \subsection*{Carol Gilligan's Critique of Kohlberg's Theory}
    \begin{itemize}
        \item Carol Gilligan argued that woman and ment simply \textbf{evaluate ethical dilemmas differently}. \textbf{Men} tend to view ethical dilemmas \textbf{legalistically} by appeal to universal moral rules, \textbf{women} tend to view ethical dilemmas \textbf{in terms of the concepts of care and concern}.
        \item She believes that \textbf{both ethical perspectives} have \textbf{something to offer}.
        \item She also argued that in \textbf{terms of ethical evaluations}, \textbf{women's opinions} are often \textbf{undervalued and ignored}.
        \item The following are some \textbf{dichotomies} between \textbf{male} and \textbf{female} approaches to ethics, alleged by Gilligan:
        \begin{enumerate}
            \item Men employ abstract moral principles in making moral evaluations, whereas women do not appeal to general principles.
            \begin{itemize}
                \item George Sher criticized this by saying that Women also employ moral rules in making decisions, but their rules are more nuanced than the rules employed by men.
            \end{itemize}
            \item Men make their moral decisions using abstract principles without reference to context, whereas women make their moral decisions by appealing to the contextual features of a given situation.
            \begin{itemize}
                \item George Sher criticized this by saying that even abstract rules cannot be applied unless the context of the situation is known.
            \end{itemize}
            \item Men rely on te concepts of justice and rules, whereas women rely on the concepts of care and concern.
            \begin{itemize}
                \item George Sher criticized this by saying t hat moral rules are not inconsistent with care and concern, and for that matter, care also involves consideration of quality, justice, and fairness.
            \end{itemize} 
        \end{enumerate}
        \item 
    \end{itemize}

    \section*{The Ethics of Care}

    \subsection*{Virginia Held's Ethics of Care (Feminist Ethics)}
    \begin{itemize}
        \item Virginia Held argues that \textbf{the ethics of care} has the following important features:
        \begin{enumerate}
            \item The \textbf{central focus} of a \textbf{care ethic} is attending to the \textbf{needs of others}, especially with respect to whom we are \textbf{responsible for} (family, friends, etc).
            \item The \textbf{ethics of care} \textbf{condones} the use of \textbf{some emotions} in making \textbf{ethical decisions}, but not all emotions (such as envy, greed, etc).
            \item The \textbf{ethics of care rejects} the use of \textbf{abstract reasoning} in making \textbf{ethical decisions}.
            \item The \textbf{care of ethics} re-conceptualizes traditional notions about \textbf{the public and the private} (for example: women stay home in privagte, while men go out and work in public).
            \item The \textbf{ethics of care} uses the \textbf{concept of a person} as a \textbf{relation being} as opposed to being a \textbf{self-sufficient, stand-alone rational agent}.
        \end{enumerate}
        \item Held notes that \textbf{traditional moral theories} portray \textbf{moral dilemmas} as involving a \textbf{conflict} between \textbf{egoistic concerns} and \textbf{universal moral principles}. This is something that is \textbf{omitted} from \textbf{the ethics of care}.
        \item She also notes that \textbf{traditional moral theories} that ignore the \textbf{domain on the household and the family} leads to \textbf{unequal power dynamics} that \textbf{women and children do not choose}.
    \end{itemize}

\end{document}